Parallel algorithms for one-dimensional multiextremal optimization

### 2017-10-06 2017/ParOptBook/Chapter3.tex
Added items remotely
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter4.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter5.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter6.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter7.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\Chapter8.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\dedic.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\editor.tex
D:\Common\Barkalov\Публикации\2016 ParOptBook Eng\LaTeX\foreword.tex
D:\Common\Barkalov\Публикации\20

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Name of First Author and Name of Second Author}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Name of First Author \at Name, Address of Institute, \email{name@email.address}
\and Name of Second Author \at Name, Address of Institute \email{name@email.address}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content. The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access. This allows unregistered users to read the abstract as a teaser for the complete chapter. As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf. source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript. Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}

\section{Section Heading}
Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.

Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations. And please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

>>>>

%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\documentclass[graybox]{svmult}
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%\usepackage{multirow}
%\usepackage{array}
%\usepackage{caption}
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

\title{Parallel algorithms for one-dimensional multiextremal optimization}
%\titlerunning{Global optimization via dimensionality reduction}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Vladimir A.Grishagin and Yaroslav D.Sergeyev}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Vladimir A.Grishagin \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia   \email{vagris@unn.ru}
\and Yaroslav D.Sergeyev \at N.I.Lobachevsky State University, Gagarin Avenue 23, 603950 Nizhni Novgorod, Russia 
\at DIMES, University of Calabria, Via P. Bucci, Cubo 42-C, 87036 Rende (CS), Italy \email{yaro@si.deis.unical.it}}
%
\maketitle

\abstract{This chapter is devoted to the development and study of parallel global search algorithms on the base of a generalization of the sequential characteristical scheme to the case of parallel computations.  Classes of synchronous and asynchronous optimization algorithms are introduced following the general schemes of parallel computations for solving optimization problems described in Chapter 1. The problems of convergence and parallelization efficiency for these characteristical algorithms are theoretically investigated including a study of various types of convergence, sufficient conditions of convergence to global minimizers, and conditions of non-redundant parallelization. The general theoretical results are applied to a number of known global optimization methods and are illustrated in computational experiments.}

\section{Parallel synchronous characteristical algorithms of global search}
\label{sec:3_1}
Let us consider the one-dimensional problem 
\begin{equation}
\label{eq:3_1}
\varphi(x)\rightarrow \min,\ x\in Q=[a,b]
\end{equation}

for the search domain  $Q=[a,b]$ and the general schemes of the parallel synchronous optimization methods (\ref{eq:1_20})-(\ref{eq:1_25}) and the parallel asynchronous methods (\ref{eq:1_26})-(\ref{eq:1_30}). Let us assume that the termination condition is absent in these schemes, i.e., in (\ref{eq:1_25}) $H_{k(n)}(\bullet) = 1$ and in (\ref{eq:1_30}) $H_{k}(\bullet)=1$ for any $k=k(n)$ will be the total number of trials executed during all $n$ iterations.

\begin{definition}
	A parallel method for solving the problem (\ref{eq:3_1}) is called \textit{parallel synchronous characteristical algorithm}, if, starting from some iteration number $n_0\ge1$, the selection of the trial coordinates at the $(n+1)$-th iteration $(n\ge n_0)$ is performed according to the following procedure:

	\begin{description} [Step 1.]
	\item [\textbf{Step 1.}] {Define a set 
	\begin{equation}
	\label{eq:3_2}
	\Lambda_n = \{x_0,x_1,...,x_\tau \}
	\end{equation}
	of a finite number $\tau + 1 = \tau(n)+1$ of the points in the domain $Q=[a,b]$ assuming that $a \in \Lambda_n, b \in \Lambda_n$, all coordinates of the preceding trials $x^i \in \Lambda_n, 1 \le i \le k$, and the set $\Lambda_n$ is ordered (by the subscript) in the increasing order of the coordinates, i.e.,
	\begin{equation}
	\label{eq:3_3}
	a = x_0 < x_1 < ... <x_{\tau - 1} < x_\tau = b.
	\end{equation}
	}
	\item [\textbf{Step 2.}] {Juxtapose the value $R(i)$ called the characteristic of the interval to each interval $(x_{i-1}, x_i), 1 \le i \le \tau$.}

	\item [\textbf{Step 3.}] {Sort the characteristics $R(i), 1 \le i \le \tau$, in the decreasing order:
	\begin{equation}
	\label{eq:3_4}
	R(t_1) \ge R(t_2) \ge R(t_{\tau - 1}) \ge R(t_\tau).
	\end{equation}
	}
	\item [\textbf{Step 4.}] {Select $p = p(n+1) \le \tau$ first characteristics from the series (3.4) with the numbers $t_j, 1 \le j \le p$, and to choose the new trials of the next $(n+1)_{th}$ iteration at the points of the set 
	\begin{equation}
	\label{eq:3_5}
	T(n+1) = \{x^{k+1}, ...,x^{k+p}\},
	\end{equation}
	according to the rule
	\begin{equation}
	\label{eq:3_6}
	x^{k+j} = d(t_j) \in (x_{t_j - 1}, x_{t_j})
	\end{equation}
	for $1 \le j \le p$, in the intervals with maximal characteristics.}
	\end{description}
\end{definition}

As the examples of the parallel characteristical algorithms one can consider the parallel versions of the scanning method, of the  method of broken lines, of Kushner’s method, and of the information algorithms, the sequential variants of which have been described in the previous Chapter.

The class of the parallel characteristical methods has been introduced and investigated in (\cite{3_GrishaginSergeyevStrongin}). In this work, the succession of the convergence properties of the parallel characteristical methods with respect to the sequential prototypes has been established. Also, the estimates of the speed up of the parallel characteristical algorithms have been obtained. A theoretical substantiation of the properties of the synchronous parallel characteristic methods with respect to the character of their convergence and the parallelization efficiency will be provided below.

\section{Convergence of characteristical algorithms}
\label{sec:3_2}
\subsection{Bilateral convergence}
Let us introduce the notations $z_j = \varphi(x_j), 1 \le j \le \tau$, for the values of the objective function at the points $x_j \in \Lambda_k$ and introduce the values
\begin{equation*}
\xi = min \{ z_{i-1}, z_i \},
\end{equation*}
\begin{equation*}
\tilde{\xi_i} = 
\begin{cases}
\xi_i, & if\  x_{i-1} \in \{x^S\}\  and\  x_i \in \{x^S\}; \\
z_{i-1},& if\  x_{i-1} \in \{x^S\}\  and \ x_i \notin \{x^S\}; \\
z_i, & if\  x_{i-1} \notin \{x^S\}\  and\  x_i \in \{x^S\}.
\end{cases}
\end{equation*}

\begin{theorem}
\label{theor:3_1}
	Let a point $x_\ast$ be a limit point (or an accumulation point) of the search trials sequence $\{x_k\}$  generated by a synchronous characteristical algorithm when solving the problem (3.1) over the interval $[a, b]$, and $x_\ast \neq a$  and $x_\ast \neq b $. Assume that the characteristics $R(i)$ and the rules for the selection of the next trial points $d(t)$ satisfy the following requirements:

	\textbf{i)} if a point $\bar{x} \in [x_{i(n)-1}, x_{i(n)}]$ and $x_{i(n)-1} \to \bar{x}, x_{i(n)} \to \bar{x}$ when $n \to \infty $  then
	\begin{equation}
	\label{eq:3_7}
	R(i(n)) \to - \mu \varphi(\bar{x}) + c;
	\end{equation}

	\textbf{ii)} in the case, when starting from some search iteration, the interval $(x_{i-1}, x_i), i=i(n)$, does not include trial points, i.e., a number $\tilde{n} \ge 1$ exists such that for any $n \ge \tilde{n}$
	\begin{equation}
	\label{eq:3_8}
	(x_{i-1} ,x_{i} )\bigcap \{ x^{k} \} =\emptyset
	\end{equation}
	for the characteristic of the interval holds the strict inequality
	\begin{equation}
	\label{eq:3_9}
	\mathop{\lim }\limits_{n\to \infty } R(i)>-\mu \tilde{\xi }_{i} +c
	\end{equation}

	\textbf{iii)} when selecting a point for the next trial from set (2.59) the inequality 
	\begin{equation}
	\label{eq:3_10}
	{\rm \; \; }\max \{ x^{k+j} -x_{t_{j} -1} ,x_{t_{j} } -x^{k+j} \} \le \nu (x_{t_{j} } -x_{t_{j} -1} )
	\end{equation}
	takes place where $\mu$, $c$, and $\nu$ are some constants, and $\mu \ge 0,{\rm \; \; }0<\nu <1$.

	Then the sequence $\{ x^{k} \} $ includes two subsequences; one of those converges to $x^{*} $ from the left, and another one from the right.
	\end{theorem}
	\begin{proof} Let us first consider the case when $x^{*} \notin \{ x^{k} \} $. Let us denote as $s=s(n),n\ge 1$, the number of the interval $(x_{s-1} ,x_{s} )$ including the limit point $x^{*} $ after the execution of the $n$-th iteration. Obviously, after the first iteration $[x_{s-1} ,x_{s} ]\subseteq [a,b]$. After the next trial point $x^{k+l} \in T(n+1)$ has fallen into the interval $(x_{s-1} ,x_{s} )$, the following estimate holds according to (3.10) for a new interval $(x_{s(n+1)-1} ,x_{s(n+1)} )$ containing $x^{*} $:

	\begin{equation*}
		x_{s(n+1)} -x_{s(n+1)-1} \le \nu (x_{s(n)} -x_{s(n)-1} )
	\end{equation*}

	In this case, if $\lambda$ trials (starting from the beginning of the search process) fell into the interval including the point $x^{*} $, the length of this interval satisfies the inequality 
	\begin{equation}
	\label{eq:3_11}
	x_{s} -x_{s-1} \le \nu ^{\lambda } (b-a)
	\end{equation}

	Since the point $x^{*}$ is a limit one, after the formation of the interval $(x_{s-1} ,x_{s} )$ at some iteration, an infinite number of trials falls into this interval. Therefore, it follows from (\ref{eq:3_11}) that 
	\begin{equation}
	\label{eq:3_12}
	\mathop{\lim }\limits_{n\to \infty } (x_{s(n)} -x_{s(n)-1} )=0 
	\end{equation}

	According to  (\ref{eq:3_12}), we can take the sequence of the left ends $\{ x_{s(n)-1} \} $ and the one of the right ends $\{ x_{s(n)} \} $ of intervals $(x_{s(n)} ,x_{s(n)-1} )$ as the sought subsequences.

	If $x^{*} \in \{ x^{k} \} $, the numbers $w\ge 1$ and $q>k(w-1)$ such that $x^{q} =x^{*} \in T(w)$ can be always found. Then for any $n>w$ a number $j=j(n)$, $0\le j\le \tau (n)$, exists, for which $x_{j} =x^{*} $. Assume that a unilateral convergence to $x^{*} $ takes place, for example, from the left (the case of the unilateral convergence from the right can be considered in the same way). Then an iteration number $\tilde{w}\ge w$ and the trial number $v>k(\tilde{w})$ can be always found so that $x^{v} \in T(\tilde{w}),$ and for all $n\ge \tilde{w}$ the trials will not fall into the interval $(x_{j} ,x_{j+1} )=(x^{q} ,x^{v} )$.

	For the interval $(x_{j} ,x_{j+1} )$ according to  (\ref{eq:3_8}),  (\ref{eq:3_9})

	\begin{equation*}
	\mathop{\lim }\limits_{n\to \infty } R(j+1)>-\mu \, \tilde{\xi }_{j+1} +c\ge -\mu \, \varphi (x^{*} )+c.
	\end{equation*}

	Correspondingly, for the interval $(x_{j-1} ,x_{j} )$ because of  (\ref{eq:3_7})

	\begin{equation}
	\label{eq:3_13}
	\mathop{\lim }\limits_{n\to \infty } R(j)=-\mu \, \varphi (x^{*} )+c.
	\end{equation}

	Therefore, starting from some iteration, the inequality $R(j+1)>R(j)$ will hold.

	However, because of the decision rules  (\ref{eq:3_4}) - (\ref{eq:3_6}) this relation contradicts to the impossibility to execute the trials in the interval $(x_{j} ,x_{j+1} )$. 
	\end{proof}
	\begin{corollary} 
	A termination condition of the kind 
		\begin{equation}
	\label{eq:3_14}
	\mathop{\min }\limits_{1\le j\le p} (x_{t_{j} } -x_{t_{j} -1} )\le \varepsilon ,
	\end{equation}
	where the numbers $t_{j} ,\; 1\le j\le p$, are from  (\ref{eq:3_4}) and $\varepsilon >0$ is the predefined search precision can be introduced into the computational scheme of the characteristical algorithm satisfying the conditions of Theorem \ref{theor:3_1}. In other words, the computations are terminated as soon as the length of the smallest interval among the intervals of the iteration with the maximal characteristics becomes less than the predefined precision $\epsilon$. In this case, the search process is terminated after a finite number of steps. 	
	\end{corollary}
	\begin{proof} First, note that in a finite interval $[a,b]$ the trial sequence will always have at least one limiting point $x^{*} $. Let us denote as $s=s(k),k\ge 1$, the number of the interval including the point $x^{*} $ at the $n$-th iteration of the search. Since this point is a limit one, an infinite number of trials falls into the interval $(x_{s-1} ,x_{s} )$. In this case, relation  (\ref{eq:3_12}) takes place for this interval, from  where it follows that condition  (\ref{eq:3_14}) is fulfilled inevitably at some search iteration. Note that if the point $x^{*} $ is not an internal one, a unilateral convergence is sufficient for feasibility of  (\ref{eq:3_14}).
	\end{proof}
	\textbf{Note.} Condition  (\ref{eq:3_10}) is equivalent to the inequality
	\begin{equation*}
	{\rm \; \; }\min \{ x^{k+j} -x_{t_{j} -1} ,x_{t_{j} } -x^{k+j} \} \ge (1-\nu )(x_{t_{j} } -x_{t_{j} -1} ).
	\end{equation*}

	Indeed, one can see easily that, for example,
	\begin{displaymath}
	x^{k+j} -x_{t_{j} -1} =(x_{t_{j} } -x_{t_{j} -1} )-(x_{t_{j} } -x^{k+j} )\ge (x_{t_{j} } -x_{t_{j} -1} )-\nu (x_{t_{j} } -x_{t_{j} -1} )=  
	\end{displaymath}
	\begin{displaymath}
	=(1-\nu )(x_{t_{j} } -x_{t_{j} -1} ).  
	\end{displaymath}

	For the interval $x_{t_{j} } -x^{k+j} $, an analogous estimate takes place, from were it follows that the considered inequality is true.

	Let us verify the feasibility of the conditions of Theorem  \ref{theor:3_1} for several examples of the algorithms belonging to the caharacteristical class. These algorithms are generalizations of known sequential methods of global optimization (see Chapter 2) to the parallel case. \\

	\textbf{Method of sequential scanning } \\

	For this method the characteristics 
	\[R(i)=x_{i} -x_{i-1} ,\, 1\le i\le \tau ,\] 
	and the new point  $x^{k+j} ,\; 1\le j\le p,$ from  (\ref{eq:3_6}) in the interval $(x_{t_{j} -1} ,x_{t_{j} } )$is calculated as
	\[x^{k+j} =0.5(x_{t_{j} -1} +x_{t_{j} } ).\]

	If the interval $(x_{i-1} ,x_{i} )$ shrinks into a point, the characteristic of this method tends to zero. Therefore, one can take the constants $\mu$ and $c$ in  (\ref{eq:3_7}) as $\mu = c = 0 $. If the trials do not fall into the interval $(x_{i-1} ,x_{i} )$ starting from some search iteration, the length of this interval (coinsiding with its characteristic) remains positive that provides the fulfillment of the condition  (\ref{eq:3_9}) for selected $\mu$ and $c$. As for inequality  (\ref{eq:3_10}), it obviously is true for  $\nu = 0.5$.
	\\

	\textbf{Kushner's method }\cite{3_Kushner} \\

	In this algorithm the characteristics 
	\[R(i)=-\frac{4(\varphi _{k}^{*} -\delta _{k} -z_{i} )(\varphi _{k}^{*} -\delta _{k} -z_{i-1} )}{x_{i} -x_{i-1} } ,\; 1\le i\le \tau ,\] 
	are used, and the new trial  (\ref{eq:3_6}) in the interval  $(x_{t-1} ,x_{t} ),\, t=t_{j} ,$  are executed at the point
	\begin{equation} 
	\label{eq:3_15} 
	x^{k+j} =x_{t-1} +\frac{(x_{t} -x_{t-1} )(\varphi _{k}^{*} -\delta _{k} -z_{t} )}{2(\varphi _{k}^{*} -\delta _{k} )-z_{t} -z_{t-1} }  
	\end{equation} 
	where $\delta _{k} >0$ is a parameter of the method and 
	\[\varphi _{k}^{*} =\mathop{\min }\limits_{0\le i\le \tau } z_{i} .\]

	Let us assume that the function \textit{$\varphi$}(\textit{x}) is limited within the finite values $\varphi _{\min } $ and $\varphi _{\max } $ over the interval $[a,b]$, i.e., $\varphi _{\min } \le \varphi (x)\le \varphi _{\max } ,\; x\in [a,\, b]$.

	If for any step $k$ the method parameter $\delta _{k} >\delta >0$ then $\varphi _{k}^{*} -z_{i} -\delta _{k} <-\delta <0$, $\varphi _{k}^{*} -z_{i-1} -\delta _{k} <-\delta <0$ and the characteristic 
	\[R(i(k))=-\frac{4(\varphi _{k}^{*} -\delta _{k} -z_{i} )(\varphi _{k}^{*} -\delta _{k} -z_{i-1} )}{x_{i} -x_{i-1} } \to -\infty \] 
	when the length of the interval $(x_{i-1} ,x_{i} )$ tends to zero. Therefore,  (\ref{eq:3_7}) holds when $\mu = 0$ and the nominal value of $c=-\infty $ (the proof of Theorem \ref{theor:3_1} does not change if the limit value in the right part of  (\ref{eq:3_7}) is equal to $-\infty$).

	Since the characteristic of any interval with a nonzero length possesses a finite value,  (\ref{eq:3_9}) holds as well.

	Let us remind a notation $t=t_{j} $ for the trial $x^{k+j} $of the current iteration  (\ref{eq:3_15}) and consider the differences

	\begin{equation*}
	x^{k+j} -x_{t-1} =\beta _{1} (x_{t} -x_{t-1} ),\; \; x_{t} -x^{k+j} =\beta _{2} (x_{t} -x_{t-1} ), 
	\end{equation*}

	where

	\begin{equation*}
	\beta _{1} =\frac{z_{t-1} -\varphi _{k}^{*} +\delta _{k} }{z_{t-1} +z_{t} -2(\varphi _{k}^{*} -\delta _{k} )} ,\; \; \beta _{2} =\frac{z_{t} -\varphi _{k}^{*} +\delta _{k} }{z_{t-1} +z_{t} -2(\varphi _{k}^{*} -\delta _{k} )}.
	\end{equation*}

	Let us introduce an auxiliary function $\sigma (w,\alpha )=\frac{w+\alpha }{w+2\alpha } $. Since $\min \{ z_{t-1} ,\; z_{t} \} \ge \varphi _{k}^{*} $, $\beta _{1} \le \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta _{k} ),\; \beta _{2} \le \sigma (z_{t} -\varphi _{k}^{*} ,\delta _{k} )$.

	For any positive value $w$, the function $\sigma (w,\alpha )$ decreases with increasing $\alpha$ because $\sigma '_{\alpha } (w,\alpha )=\frac{-w}{(w+2\alpha )^{2} } <0,$ therefore, for $\delta _{k} >\delta >0$ the following inequalities hold:

	\begin{equation*}
	\beta _{1} \le \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta ),\; \beta _{2} \le \sigma (z_{t} -\varphi _{k}^{*} ,\delta ).
	\end{equation*}

	In turn, the function $\sigma (w,\alpha )$ increases with increasing $w$ for any positive $\alpha$ since $\sigma '_{w} (w,\alpha )=\frac{\alpha }{(w+2\alpha )^{2} } >0$, therefore

	\begin{equation*}
	\max \{ \sigma (z_{t-1} -\varphi _{k}^{*} ,\delta ),\; \sigma (z_{t} -\varphi _{k}^{*} ,\delta )\} \le \sigma (\varphi _{\max } -\varphi _{\min } ,\; \alpha ). 
	\end{equation*}

	The last inequality means that

	\begin{equation*}
	\max \{ \beta _{1} ,\; \beta _{2} \} \le \sigma (\varphi _{\max } -\varphi _{\min } ,\; \delta )=\frac{\varphi _{\max } -\varphi _{\min } +\delta }{\varphi _{\max } -\varphi _{\min } +2\delta }. 
	\end{equation*}

	The last value can be selected as the constant $\nu$  for the relation  (\ref{eq:3_10}) since, obviously, $0<\nu <1$. \\

	\textbf{Method of broken lines \cite{3_Piyavskij, 3_Shubert}} \\

	The characteristics of this algorithm

	\begin{equation}
	\label{eq:3_16}
	R(i)=\frac{m(x_{i} -x_{i-1} )}{2} -\frac{z_{i} +z_{i-1} }{2} ,\; 1\le i\le \tau,
	\end{equation}
	and for the interval $(x_{t_{j} -1} ,x_{t_{j} } )$ from  (\ref{eq:3_6}) the point  of the new trial falling into this interval is taken in the form

	\begin{equation}
	\label{eq:3_17}
	x^{k+j} =\frac{x_{t_{j} } +x_{t_{j} -1} }{2} -\frac{z_{t_{j} } -z_{t_{j} -1} }{2m}.
	\end{equation}

	In order to satisfy the assertion of the theorem, let us assume that the minimized function $\varphi(x)$ satisfies the Lipschitz condition with the constant $L>0$, i.e.,

	\begin{equation}
	\label{eq:3_18}
	\left|\varphi (x')-\varphi (x'')\right|\le L\left|x'-x''\right|,x',x''\in [a,b],
	\end{equation}	
	and, in addition, the method parameter $m>L$.

	Being Lipschitzian, the function $\varphi(x)$ is continuous. Hence, the characteristic of the interval will tend to the value $-\varphi (\bar{x})$ when shrinking the interval $(x_{i-1} ,x_{i} )$ to the point $\bar{x}$, i.e., for the feasibility of  (\ref{eq:3_7}) one can take $\mu= 1$ and $c=0$. Now let us check if inequality  (\ref{eq:3_9}) holds with given $\mu$ and $c$ for the interval $(x_{i-1} ,x_{i} )$ satisfying  (\ref{eq:3_8}). We can use a simple equality

	\begin{equation}
	\label{eq:3_19}
	\min \{ z_{i-1} ,z_{i} \} =\frac{1}{2} (z_{i-1} +z_{i} -\left|z_{i-1} -z_{i} \right|)
	\end{equation}
	to estimate the method characteristic  (\ref{eq:3_16}), along with the Lipschitz condition and the assumption  (\ref{eq:3_8}) that the length of the interval remains constant and positive, starting from some search iteration:
	\begin{displaymath}
	R(i)=m\frac{x_{i} -x_{i-1} }{2} -\frac{z_{i} +z_{i-1} }{2} >L\frac{x_{i} -x_{i-1} }{2} -\frac{z_{i} +z_{i-1} }{2} \ge \frac{\left|z_{i-1} -z_{i} \right|-(z_{i-1} +z_{i} )}{2} =
	\end{displaymath}
	\begin{displaymath}
	=-\min \{ z_{i-1} ,z_{i} \}. 
	\end{displaymath}

	The obtained inequalities establish the truth of  (\ref{eq:3_9}).

	In order to determine the factor $\nu $ in  (\ref{eq:3_10}) let us estimate the value
	\begin{displaymath}
	x_{t_{j} } -x^{k+j} =\frac{x_{t_{j} } -x_{t_{j} -1} }{2} +\frac{z_{t_{j} } -z_{t_{j} -1} }{2m} \le \frac{x_{t_{j} } -x_{t_{j} -1} }{2} +L\frac{x_{t_{j} } -x_{t_{j} -1} }{2m} = 
	\end{displaymath}
		\begin{displaymath}
	 =\left(1+\frac{L}{m} \right)\frac{x_{t_{j} } -x_{t_{j} -1} }{2} . 
	\end{displaymath}

	An analogous estimate takes place for the interval $(x_{t_{j} -1} ,x^{k+j} )$ as well, therefore, one can take $\nu =0,5(1+L/m)$. Obviously, $\nu >0$ and, because $m>L$, $\nu <1$. \\

	\textbf{Core information Algorithm of Global Search (AGS)} \\

	The decision rule  (\ref{eq:3_2})- (\ref{eq:3_6}) of AGS includes characteristics

	\begin{equation}
	\label{eq:3_20}
	R(i)=m(x_{i} -x_{i-1} )+\frac{(z_{i} -z_{i-1} )^{2} }{m(x_{i} -x_{i-1} )} -2(z_{i} +z_{i-1} ),\; 1\le i\le \tau ,
	\end{equation}
	and points of new trials calculated according to  (\ref{eq:3_17}). The factor $m$ is determined by the expressions

	\begin{equation}
	\label{eq:3_21}
	m=\left\{\begin{array}{c} {rM,M>0} \\ {1,M=0} \end{array}\right. ,
	\end{equation}
	\begin{equation}
	\label{eq:3_22}
	M=\mathop{\max }\limits_{1\le i\le \tau } \frac{\left|z_{i} -z_{i-1} \right|}{x_{i} -x_{i-1} },
	\end{equation}
	where $r>1$ is a parameter of the method.

	Having assumed the Lipschitz property  (\ref{eq:3_18}) of the objective function  (\ref{eq:3_1}), let us prove that this method also satisfies the conditions of Theorem \ref{theor:3_1} with $\mu=4$, $c=0$, and $\nu =0,5(1+1/r)$.

	Because of Lipschitzness when shrinking the interval $(x_{i-1} ,x_{i} )$ into a point $\bar{x}$ its characteristic tends to $-4\varphi (\bar{x})$, i.e., for the truth of  (\ref{eq:3_7}) one can take $\mu=4$ and $c=0$.

	In order to check  (\ref{eq:3_9}) under the condition  (\ref{eq:3_8}) let us consider two cases. First, assume that the equality $z_{i-1} =z_{i} $ holds for the interval $(x_{i-1} ,x_{i} )$. Then, the characteristic
	\begin{equation*}
	R(i)=m(x_{i} -x_{i-1} )-2(z_{i} +z_{i-1} )>\; -2(z_{i} +z_{i-1} )= -4\min \{ z_{i-1} ,z_{i} \},
	\end{equation*}
	since the length of the interval $(x_{i-1} ,x_{i} )$ does not change anymore, starting from some iteration $n_{\Delta } $ , i.e., a constant $\Delta >0$ exists such that for all $n>n_{\Delta } $ the inequalities $x_{i} -x_{i-1} >\Delta >0$ hold.

	Now assume that $z_{i-1} \ne z_{i} $. Let us rewrite the characteristic  (\ref{eq:3_20}) in the form

	\begin{equation}
	\label{eq:3_23}
	R(i) = |z_{i} -z_{i-1} |(\beta +\frac{1}{\beta } )-2(z_{i} +z_{i-1} ),
	\end{equation}
	where $0<\beta =\frac{m(x_{i} -x_{i-1} )}{\; \left|z_{i} -z_{i-1} \right|} \; <1$ as a consequence of  (\ref{eq:3_21}),  (\ref{eq:3_22}). In this case, the value $\beta +\frac{1}{\beta } >2$, therefore from  (\ref{eq:3_23})

	\begin{equation*}
	R(i)>\; 2\left|z_{i} -z_{i-1} \right|-2(z_{i} +z_{i-1} )=-4\min \{ z_{i} ,z_{i-1} \} .
	\end{equation*}

	As for the condition  (\ref{eq:3_10}), according to  (\ref{eq:3_22}) the following estimate holds: 
	\begin{displaymath}
	x_{t_{j} } -x^{k+j} =\frac{x_{t_{j} } -x_{t_{j} -1} }{2} +\frac{z_{t_{j} } -z_{t_{j} -1} }{2m} \le \frac{x_{t_{j} } -x_{t_{j} -1} }{2} +M\frac{x_{t_{j} } -x_{t_{j} -1} }{2m} =
	\end{displaymath}
	\begin{displaymath}
	=\left(1+\frac{1}{r} \right)\frac{x_{t_{j} } -x_{t_{j} -1} }{2}. 
	\end{displaymath}

	The length of the interval $(x_{t_{j} -1} ,x^{k+j} )$ can be estimated in the same way, therefore, one can take $\nu$ = $0.5(1+1/r)$. Since $r>1$, the condition $0<\nu <1$ is satisfied obviously. \\

	\textbf{Information Algorithm of Global Search with trials at internal points (AGSIP)} \\

	In the framework of this method all trials are executed at internal points of the search domain $[a,b]$. The set $\Lambda _{n} $ includes the trial points and the end points $a$ and $b$, from where $\tau =k+1$. Morover, $z_{j} =\varphi (x_{j} )$ for $1\le j\le k$ only since at the ends points of the domain the trials are not executed.

	The characteristics $R(i),\; 1\le i\le \tau ,$ of AGSIP are computed in accordance with the expression

	\begin{equation}
	\label{eq:3_24}
	R(i)=
	\left\{
	\begin{array}{l} 
	{m(x_{i} -x_{i-1} )+\frac{r(z_{i} -z_{i-1} )^{2} }{M(x_{i} -x_{i-1} )} -2(z_{i} +z_{i-1} ),\; 1<i<\tau ,} \\ 
	{m(x_{i} -x_{i-1} )-4z_{1} ,\; i=1,} \\ 
	{m(x_{i} -x_{i-1} )-4z_{k} ,\; i=\tau ,} 
	\end{array}
	\right., 
	\end{equation}

	and the next trial points are computed as

	\begin{equation}
	\label{eq:3_25}
	x^{k+j} =\; 0.5(x_{t_{j} } +x_{t_{j} -1} )-
	\left\{
	\begin{array}{l} 
		{0,\; \; t_{j} =1\; \; or \; \; t_{j} =\tau ,} \\ 
		{(z_{t_{j} } -z_{t_{j} -1} )/(2m),\; 2\le t_{j} \le \tau -1.} 
	\end{array}
	\right. 
	\end{equation}

	Under Lipschitz condition  (\ref{eq:3_18}), we obtain that when shrinking the interval $(x_{i-1} ,x_{i} )$ to a point $\bar{x}$, its characteristic tends to $-4\varphi (\bar{x})$, i.e., for the fulfillment of  (\ref{eq:3_7}), it is possible to take $\mu = 4$ and $c = 0$.
